Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-24 NEUROC/Response-MajRev/response_letter.tex

% !TEX program=pdflatex
% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{times}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{multirow}
\usepackage{lscape}
\usepackage{xcolor}
\usepackage{float}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{framed}
\usepackage[colorlinks=true]{hyperref}
% Vector
\newcommand{\MyVec}[1]{\boldsymbol{#1}}
% Matrix
\newcommand{\MyMat}[1]{\boldsymbol{#1}}
% Vectors with hat
\newcommand{\hvec}[1]{\hat{\boldsymbol{h}}_{#1}}
% Model parameters
\newcommand{\Param}[1]{\textcolor{blue}{#1}}

\usepackage{blindtext}
\usepackage[most]{tcolorbox} 
\definecolor{block-gray}{gray}{0.95}

\newtcolorbox{zitat}[2][]{%
    colback=block-gray,
    grow to right by=-10mm,
    grow to left by=-10mm, 
    boxrule=0pt,
    boxsep=0pt,
    breakable,
    enhanced jigsaw,
    borderline west={4pt}{0pt}{gray},
    title={#2\par},
    colbacktitle={block-gray},
    coltitle={black},
    fonttitle={\large\bfseries},
    attach title to upper={},
    #1,
}

\newenvironment{myquote}[1]%
{\vspace{0.5em}\begin{zitat}{#1}}
{\end{zitat}\vspace{0.5em}}

@@

    In the revised manuscript, we use $e_{x,y}$ to represent an edge and use $\boldsymbol{e}_{x,y}$ to represent its input feature vector.
    %
    The input feature vectors of all edges are same for all GNN layers.
    %
    We use $\boldsymbol{s}$ to represent aggregated vectors outputted by the aggregation function $\Sigma$ in graph neurons.
    %
    For every vertex $v_x$, we use $\boldsymbol{s}^l_x$ to denote its aggregated vector in the GNN layer $l$.
    %
    If the GNN layer $l$ has sub-layers, $\boldsymbol{s}^{l,i}_x$ represents its aggregated vector in the $i$-th sub-layer.

    To clarify the concept of \emph{sub-layers} in a GNN layer, we have added more description on it in the revised manuscript.
    %
    We first introduce the concept of sub-layer in Section 2.2 ``Graph Neuron and Message-passing Model'' as:
    %
    \begin{myquote}{Section 2.2}
        Some complex GNNs like GAT [6] and GaAN [7] use more than one message passing phase in each GNN layer.
        %
        We regard every message passing phase in a GNN layer as a \emph{sub-layer}.
        %
        We will give out more details on sub-layers when we introduce GAT.
    \end{myquote}
    %
    We then use GAT as an example to elaborate on the concept of sub-layers in Section 2.3 ``Representative GNNs'' as:
    %
    \begin{myquote}{Section 2.3}
        \newcommand{\GATCalcWeight}{\exp(LeakyReLU(\Param{\MyVec{a}}^T[\hvec{y}[k] \parallel \hvec{x}[k]]))}
        Each GAT layer consists of a vertex pre-processing phase and two sub-layers (i.e., message-passing phases).

@@

% R1Q2
\begin{point}
    Some typical applications of GNNs should be included, such as video object segmentation [ref1], human-object interaction [ref2] and human-parsing [ref3].[1] Zero-shot video object segmentation via attentive graph neural networks,iccv 2019 [2] Learning human-object interactions by graph parsing neural networks, eccv 2018. [3] Hierarchical human parsing with typed part-relation reasoning, cvpr 2020.
\end{point}

\begin{reply}
	Thank you for pointing out our shortcomings. Computer vision is indeed another important application area of graph neural networks. We have added the mentioned references in the INTRODUCTION section in the revised manuscript. We quote the related sentence below:
    \begin{myquote}{Section 1}
         The powerful expression ability makes GNNs achieve good accuracy in not only graph analytical tasks [8, 9, 10] (like node classification and link prediction) but also computer vision tasks (like human-object interaction [11], human parsing [12], and video object segmentation [13]).
    \end{myquote}
\end{reply}

@@

    Thank you for pointing the problem out.
    %
    Besides Figure 6 and Figure 7, we have adjusted all the figures in the revised manuscript to make sure that font sizes in figures are no less than the font size of figure captions.
    %
    To make the figures more easy to read, we also splitted Figure 6 and Figure 7 into five figures (Figure 6 to 10) in the revised manuscript to enlarge the subfigures.

\end{reply}

% R1Q5
\begin{point}
    In my view, computation efficiency is to describe the testing or validation process. Except for reporting and analyzing the training times, it is meaningful to discuss the inference time. This is also an important point of view for deep learning researchers to be concerned about.
\end{point}

\begin{reply}
    Thank you for the insightful comment and suggestion.
    %
    The efficiency of inference (including inference time and memory usage) is indeed important for deep learning researchers and engineers.
    %
    In the revised manuscript, we have added a new subsection (Section 2.5 ``Inference with GNNs'') in the revised manuscript to briefly review how to perform inference with GNNs.
    %
    To discuss the efficiency of GNN inference on GPUs, we have added performance bottleneck analysis on GNN inference in every subsection of Section 4 ``Evaluation Results and Analysis'':

    In Section 4.1 ``Effects of Hyper-parameters on Performance'', we have additionally measured how the inference time and the peak memory usage changed as we increased the values of the hyper-parameters.
    %
    We find that ...
    %%% TODO %%%

    In Section 4.2 ``Time Breakdown Analysis'', we have additionally conducted the time breakdown analysis for GNN inference.
    %
    We conducted the same set of experiments as in the training time anlaysis.
    %
    We found that the performance characteristics of GNN inference were very similar to GNN training.
    %
    The processing time breakdowns of GCN on the layer level, step level and operator level of training and inference were all highly similar, as shown in Figure~\ref{fig:compare_layer_level_time_breakdown} to Figure~\ref{fig:compare_basic_operators} of the response.
    %
    We observed similar phenomena in other GNNs.
    %
    Based on the time breakdown analysis, we concluded that the performance bottenecks of GNN training and inference were same.
    %
    The difference between training and inference process was mainly reflected in the elapsed time.
    %
    Given the same dataset and the same GNN, the inference time was only XX\% of the training time, as compared in Figure~\ref{} of the response.
    %
    The training process of each epoch consisted of three phases: forward propagation, backward propagation and evaluation.
    %
    The evaluation phase conducted inference for the vertices in the evaluation set.
    %
    Under the semi-supervised learning setting, the graph neurons of the whole graph participated in inference, but only the predicted labels of the veritces in the evaluation set were used to calculate accuracy.
    %
    Therefore, the elapsed time of the evaluation phase was very close to the elapsed time of inference.

    \begin{figure}[H]
        \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_layer_time_proportion_gcn.pdf}}
       % \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_layer_time_proportion_gcn.pdf}}
        \caption{Time break down on the layer level of GCN.}
        \label{fig:compare_layer_level_time_breakdown}
    \end{figure}

    \begin{figure}[H]
            \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
            %\subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
            \caption{Effects of the average degree on the time proportion of the edge/vertex calculation of GCN. Graphs were generated by fixing the number of vertices as 50,000.}
            \label{fig:compare_effect_of_degree}
     \end{figure}

    \begin{figure}[H]
            \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_edge_calc_decomposition_gcn.pdf}}
            %\subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
            \caption{Time breakdown of the edge calculation stage of GCN.}
            \label{fig:compare_time_breakdown_of_edge_calculation}
     \end{figure}

     \begin{figure}[H]
             \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_top_basic_ops_gcn.pdf}}
             %\subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_top_basic_ops_gcn.pdf}}
             \caption{Top 5 time-consuming basic operators of GCN. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
             \label{fig:compare_basic_operators}
      \end{figure}

     In Section 4.3 ``Memory Usage Analysis'', we have additionally conducted the memory usage analysis for GNN inference and added related discussion in XXX.
     %
     Figure~\ref{fig:compare_memory_expasion_ratio} in the response (Figure X in the revised manuscript) compares the memory expansion ratios of typical GNNs during training and inference.
     %
     For GCN, the memory expansion ratios (MER) of training and inference were nearly same since GCN did not cache intermediate results during training.
     %
     For GGNN, the MERs of inference were 45\% to 83\% of training.
     %
     For GAT, the MERs of inference were 52\% to 61\% of training.
     %
     For GaAN, the MERs of inference were 37\% to 69\% of training.
     %
     Though the MERs of GGNN/GAT/GaAN decreased significantly compared with training, the values of MERs were still high.
     %
     The high memory usage still prevented us inferencing on big graphs.

     \begin{figure}[H]
        \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_memory_expansion_ratio.pdf}}
        \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_memory_expansion_ratio.pdf}}
        \caption{Memory expansion ratios of typical GNNs.}
        \label{fig:compare_memory_expasion_ratio}
     \end{figure}

     %%% TODO %%%

     In Section 4.4 ``Effects of Sampling Techniques on Performance'', we have additionally evaluated the performance bottlenecks in sample-based GNN inference.
     %
     We find that ...
     %%% TODO %%%

### 2020-12-24 NEUROC/Response-MajRev/response_letter.tex

% R1Q5
\begin{point}
    In my view, computation efficiency is to describe the testing or validation process. Except for reporting and analyzing the training times, it is meaningful to discuss the inference time. This is also an important point of view for deep learning researchers to be concerned about.
\end{point}

\begin{reply}
    Thank you for the insightful comment and suggestion.
    %
    The efficiency of inference (including inference time and memory usage) is indeed important for deep learning researchers and engineers.
    %
    In the revised manuscript, we have added a new subsection (Section 2.5 ``Inference with GNNs'') in the revised manuscript to briefly review how to perform inference with GNNs.
    %
    To discuss the efficiency of GNN inference on GPUs, we have added performance bottleneck analysis on GNN inference in every subsection of Section 4 ``Evaluation Results and Analysis'':

    In Section 4.1 ``Effects of Hyper-parameters on Performance'', we have additionally measured how the inference time and the peak memory usage changed as we increased the values of the hyper-parameters.
    %
    We find that ...
    %%% TODO %%%

    In Section 4.2 ``Time Breakdown Analysis'', we have additionally conducted the time breakdown analysis for GNN inference.
    %
    We found that the performance characteristics of GNN inference were similar to GNN training.
    %
    Figure~\ref{fig:compare_layer_level_time_breakdown} of the response (Figure X in the revised manuscript) compares the time breakdowns of GCN on the layer level.
    %
    It indicates that the time breakdowns on the layer level were very similar between inference and training.
    %
    Figure~\ref{fig:compare_effect_of_degree}
    %
    The processing time breakdowns of GCN on the layer level, step level and operator level of training and inference were all highly similar, as shown in Figure~\ref{fig:compare_layer_level_time_breakdown} to Figure~\ref{fig:compare_basic_operators} of the response.
    %
    We observed similar phenomena in other GNNs.
    %
    Based on the time breakdown analysis, we concluded that the performance bottenecks of GNN training and inference were same.
    %
    The difference between training and inference process was mainly reflected in the elapsed time.
    %
    Given the same dataset and the same GNN, the inference time was only XX\% of the training time, as compared in Figure~\ref{} of the response.
    %
    The training process of each epoch consisted of three phases: forward propagation, backward propagation and evaluation.
    %
    The evaluation phase conducted inference for the vertices in the evaluation set.
    %
    Under the semi-supervised learning setting, the graph neurons of the whole graph participated in inference, but only the predicted labels of the veritces in the evaluation set were used to calculate accuracy.
    %
    Therefore, the elapsed time of the evaluation phase was very close to the elapsed time of inference.

    \begin{figure}[H]
        \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_layer_time_proportion_gcn.pdf}}
        \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_layer_time_proportion_gcn.pdf}}
        \caption{Time break down on the layer level of GCN.}
        \label{fig:compare_layer_level_time_breakdown}
    \end{figure}

@@

    \begin{figure}[H]
            \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
            %\subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
            \caption{Effects of the average degree on the time proportion of the edge/vertex calculation of GCN. Graphs were generated by fixing the number of vertices as 50,000.}
            \label{fig:compare_effect_of_degree}
     \end{figure}

    \begin{figure}[H]
            \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_edge_calc_decomposition_gcn.pdf}}
            \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_edge_calc_decomposition_gcn.pdf}}
            \caption{Time breakdown of the edge calculation stage of GCN.}
            \label{fig:compare_time_breakdown_of_edge_calculation}
     \end{figure}

     \begin{figure}[H]
             \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_top_basic_ops_gcn.pdf}}
             \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_top_basic_ops_gcn.pdf}}
             \caption{Top 5 time-consuming basic operators of GCN. The time proportion of each basic operator was averaged over all datasets with the error bar indicating the maximum and the minimum.}
             \label{fig:compare_basic_operators}
      \end{figure}

### 2020-12-24 NEUROC/Response-MajRev/response_letter.tex

    \begin{figure}[H]
        \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_layer_time_proportion_gcn.pdf}}
        \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_layer_time_proportion_gcn.pdf}}
        \caption{Time break down on the layer level of GCN.}
        \label{fig:compare_layer_level_time_breakdown}
    \end{figure}

    \begin{figure}[H]
            \subfloat[Training]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
            \subfloat[Inference]{\includegraphics[width=0.5\columnwidth]{../figs/experiments/exp_inference_full_avg_degree_on_vertex_edge_cal_time_gcn.pdf}}
            \caption{Effects of the average degree on the time proportion of the edge/vertex calculation of GCN. Graphs were generated by fixing the number of vertices as 50,000.}
            \label{fig:compare_effect_of_degree}
     \end{figure}

@@

    It is also pretty neat to divide layer time complexity into two buckets: vertex calculation and edge calculation. The data model pretty well summarizes mainstream GNN layer architectures. And this analysis is very insightful for layer profiling.

    And the experimental evaluation were done over 6 large graph-structured datasets.
\end{point}

\begin{reply}
    Thank you very much for your appreciation.	
\end{reply}

% R2Q3
\begin{point}
    While, one major drawback is that I did not clearly see the analysis complexity v.s. accuracy. For example, in Figure 19 and 20, I did not see network accuracy from those 4 GNNs. There is always tradeoff between model complexity and model performance, and in some scenarios where high complexity is allowed, a sophisticated model of more powerful representation capability is still needed.
\end{point}gg

\begin{reply}
    Thanks very much for your valuable suggestions.
    %
    The model complexity directly affected both the accuracy and the training time.
    %
    In order to analyze the relationship between model complexity and accuracy, we conducted two kinds of extra experiments in the revised manuscript: (1) how the hyper-parameters of the GNNs (like the dimension of hidden vectors and the number of heads) affected the accuracy of GNNs (in Section 4.1); (2) how the batch size in the sampling methods affected the accuracy of GNNs (in Section 4.4).

    In this reply, we focus on the first kind of experiments added in Section 4.1.
    %
    We will introduce our results of the second kind of experiments in the nextR reply.

### 2020-12-24 NEUROC/Response-MajRev/response_letter.tex

\end{reply}

% R1Q5
\begin{point}
    In my view, computation efficiency is to describe the testing or validation process. Except for reporting and analyzing the training times, it is meaningful to discuss the inference time. This is also an important point of view for deep learning researchers to be concerned about.
\end{point}

\begin{reply}
    Thank you for the insightful comment and suggestion.
    %
    The efficiency of inference (including inference time and memory usage) is indeed important for deep learning researchers and engineers.
    %
    In the revised manuscript, we have added a new subsection (Section 2.5 ``Inference with GNNs'') in the revised manuscript to briefly review how to perform inference with GNNs.
    %
    To discuss the efficiency of GNN inference on GPUs, we have added performance bottleneck analysis on GNN inference in every subsection of Section 4 ``Evaluation Results and Analysis'':

    In Section 4.1 ``Effects of Hyper-parameters on Performance'', we have additionally measured how the time and the peak memory usage changed as we increased the values of the hyper-parameters during \emph{inference}.
    %
    \begin{itemize}
    \item For all GNNs, we found that the effects of hyper-parameters on the \emph{inference time} were the same as the training time.
        %
        Taking GGNN as an example, Figure~\ref{fig:compare_effect_of_hyperparameter_on_time} in the response (Figure X and Figure X in the revised manuscript) compares the effects of the hidden vector dimension $dim(\MyVec{h}^1_x)$ on training/inference time.
        %
        As $dim(\MyVec{h}^1_x)$ increased, the vertex/edge calculation time of training time and inference time both grew linearly when the values of hyper-parameters were large enough.
        %
        We observed similar phenomena with the other GNNs.
        %
        The results indicated that the time complexity analysis of GNNs were applicable to both training and inference.

        \begin{figure}[H]
            \centering
             \subfloat[Training]{\includegraphics[width=0.7\textwidth]{../figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.pdf}}\\
            \subfloat[Inference]{\includegraphics[width=0.7\textwidth]{../figs/experiments/exp_hyperparameter_on_inference_vertex_edge_phase_time_ggnn.pdf}}
            \caption{Effects of hyper-parameters on the vertex/edge calculation time of GGNN.}
            \label{fig:compare_effect_of_hyperparameter_on_time}
        \end{figure}

        \item For all GNNs, we found that the effects of hyper-parameters on inference memory usage were also same as training.
        %
        Taking GAT as an example, Figure~\ref{fig:compare_effect_of_hyperparameter_on_memory_usage} in the response (Figure X and Figure X in the revised manuscript) compares the peak memory usage during training and inference under different values of hyper-parameters.
        %
        The trends of the curves in Figure~\ref{fig:compare_effect_of_hyperparameter_on_memory_usage_training} and Figure~\ref{fig:compare_effect_of_hyperparameter_on_memory_usage_inference} were highly similar.
        %
        The memory usage of training and inference \emph{both} grew linearly as the dimension of each head and the number of heads increased.
        %
        We observed similar phenomena with the other GNNs.
        %
        \begin{figure}[H]
            \centering
            \subfloat[Training\label{fig:compare_effect_of_hyperparameter_on_memory_usage_training}]{\includegraphics[width=0.6\textwidth]{../figs/experiments/exp_hyperparameter_on_memory_usage_gat.pdf}}\\
            \subfloat[Inference\label{fig:compare_effect_of_hyperparameter_on_memory_usage_inference}]{\includegraphics[width=0.6\textwidth]{../figs/experiments/exp_hyperparameter_on_inference_memory_usage_gat.pdf}}
            \caption{Effects of hyper-parameters on peak memory usage of GAT.}
            \label{fig:compare_effect_of_hyperparameter_on_memory_usage}
        \end{figure}

    \end{itemize}

### 2020-12-24 NEUROC/sec4-experiments.tex
更新R1P6

\paragraph{Effects on Training/Inference Time}

@@

To verify the time complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex}, we first compared the training time of the four GNNs in \figurename~\ref{fig:exp_absolute_training_time}.
%
The ranking of the training time was GaAN $\gg$ GAT $>$ GGNN $>$ GCN in all cases.
%
Since the real-world graphs had more edges than vertices ($|\mathcal{E}| > |\mathcal{V}|$), the time complexity of the edge calculation stage affected more than the vertex calculation stage.
%
The ranking was consistent with the time complexity analysis.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gcn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GCN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GGNN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gat.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GAT.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gaan.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GaAN.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan}
\end{figure}

@@

@@

We further measured the effects of the hyper-parameters on the peak GPU memory usage during training in \figurename~\ref{fig:exp_hyperparameter_memory_usage}.
%
The memory usage also increased linearly as the hyper-parameters increased for all GNNs, except for GaAN on $dim(\MyVec{h}^1_x)$.
%
As the hidden vectors $\MyVec{h}^1_x$ consumed a small proportion of memory in GaAN, the growth in the memory usage was not noticeable until $dim(\boldsymbol{h}^1_x)$ was large enough.

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[width=0.25\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[width=0.25\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_ggnn.pdf}}\\
    \subfloat[GAT]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_gat.pdf}}\\
    \subfloat[GaAN]{\includegraphics[width=0.75\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_gaan.pdf}}
    \caption{Effects of hyper-parameters on the peak GPU memory usage during training, excluding the memory used by the dataset and the model parameters.}
    \label{fig:exp_hyperparameter_memory_usage}
\end{figure}

@@

The complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} was valid.
%
Fixing other hyper-parameters, each hyper-parameter itself affected the training time and the memory usage of a GNN Layer \emph{in a linear way}.
%
Algorithm engineers could adjust hyper-parameters according to the time complexity to avoid explosive growth in the training time and memory usage.
%
The accuracy of GNNs was much more sensitive to the dimension of hidden vectors/heads than the other hyper-parameters.
%
The relative accuracy between GNNs varied greatly with different datasets.

>>>>

\paragraph{Effects on Training Time}

@@

To verify the time complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex}, we first compared the \emph{training} time of the four GNNs in \figurename~\ref{fig:exp_absolute_training_time}.
%
The ranking of the training time was GaAN $\gg$ GAT $>$ GGNN $>$ GCN in all cases.
%
Since the real-world graphs had more edges than vertices ($|\mathcal{E}| > |\mathcal{V}|$), the time complexity of the edge calculation stage affected more than the vertex calculation stage.
%
The ranking was consistent with the time complexity analysis.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gcn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GCN during training.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gcn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_ggnn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GGNN during training.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gat.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GAT during training.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{figs/experiments/exp_hyperparameter_on_vertex_edge_phase_time_gaan.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GaAN during training.}
    \label{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan}
\end{figure}

@@

\paragraph{Effects on Inference Time}

For all GNNs, we found that the effects of hyper-parameters on the inference time were the same as the training time.
%
Taking GGNN as an exmple, \figurename~\ref{fig:exp_hyperparameter_on_inference_vertex_edge_phase_time_ggnn} shows the effects of hyper-parameters on the inference time of GGNN.
%
By cross-comparing \figurename~\ref{fig:exp_hyperparameter_on_inference_vertex_edge_phase_time_ggnn} and \figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_ggnn}, the effects of hyper-parameters on inference time and training time were the same: as $dim(\MyVec{h}^1_x)$ increased, the vertex and edge calculation time both grew linearly when $dim(\MyVec{h}^1_x)$ was large enough.
%
For the other GNNs, we observed similar phenomena.
%
The results indicated that the complexity analysis presented in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} were also applicable to inference.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{figs/experiments/exp_hyperparameter_on_inference_vertex_edge_phase_time_ggnn.pdf}
    \caption{Effects of hyper-parameters on the vertex/edge calculation time of GGNN during inference.}
    \label{fig:exp_hyperparameter_on_inference_vertex_edge_phase_time_ggnn}
\end{figure}

@@

We further measured the effects of the hyper-parameters on the peak GPU memory usage during training in \figurename~\ref{fig:exp_hyperparameter_memory_usage}.
%
The memory usage also increased linearly as the hyper-parameters increased for all GNNs, except for GaAN on $dim(\MyVec{h}^1_x)$.
%
As the hidden vectors $\MyVec{h}^1_x$ consumed a small proportion of memory in GaAN, the growth in the memory usage was not noticeable until $dim(\boldsymbol{h}^1_x)$ was large enough.
%
The effects of the hyper-parameters on inference memory usage were the same as training.
%
Taking GAT as an example, the peak memory usage during inference grew linearly as the values of hyper-parameters increased, as shown in \figurename~\ref{fig:exp_inference_hyperparameter_memory_usage_gat}.
%
By cross-comparing \figurename~\ref{fig:exp_inference_hyperparameter_memory_usage_gat} and \figurename~\ref{fig:exp_hyperparameter_memory_usage_gat}, the hyper-parameters affected the peak memory usage of training and inference in the same way.
%
For the other GNNs, we observed similar phenomena.

\begin{figure}[H]
    \centering
    \subfloat[GCN]{\includegraphics[width=0.25\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_gcn.pdf}}
    \subfloat[GGNN]{\includegraphics[width=0.25\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_ggnn.pdf}}\\
    \subfloat[GAT\label{fig:exp_hyperparameter_memory_usage_gat}]{\includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_gat.pdf}}\\
    \subfloat[GaAN]{\includegraphics[width=0.75\columnwidth]{figs/experiments/exp_hyperparameter_on_memory_usage_gaan.pdf}}
    \caption{Effects of hyper-parameters on the peak GPU memory usage during training, excluding the memory used by the dataset and the model parameters.}
    \label{fig:exp_hyperparameter_memory_usage}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\columnwidth]{figs/experiments/exp_hyperparameter_on_inference_memory_usage_gat.pdf}
    \caption{Effects of hyper-parameters on the peak GPU memory usage during inference of GAT, excluding the memory used by the dataset and the model parameters.}
    \label{fig:exp_inference_hyperparameter_memory_usage_gat}
\end{figure}

@@

The complexity analysis in \tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} was valid for both training and inference.
%
Fixing other hyper-parameters, each hyper-parameter itself affected the training time and the memory usage of a GNN Layer \emph{in a linear way}.
%
Algorithm engineers could adjust hyper-parameters according to the time complexity to avoid explosive growth in the training/inference time and memory usage.
%
The accuracy of GNNs was much more sensitive to the dimension of hidden vectors/heads than the other hyper-parameters.
%
The relative accuracy between GNNs varied greatly with different datasets.
%
There was no clear winner for all datasets.
